A finite-state morphological transducer for Kyrgyz

### 2012-03-14 paper/paper.tex
minor updates

Kyrgyz has a number of nouns that underlyingly end with a consonant cluster consisting of two liquids, but which are split by an epenthetic vowel syllable-finally.
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with epenthetic vowel}\label{exrnvowel}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/мурн/ & [мурун] & \eng{nose} \\
			/мурн-LAр/ & [мурундар] & \eng{nose--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн/ & [орун] & \eng{place} \\
			/орн-LAр/ & [орундар] & \eng{places--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}
However, when a vowel-initial suffix follows the noun, the epenthetic vowel is absent, and all rules of consonant unfaithfulness involving consonant clusters are enforced[?].
%FIXME: "[?]"
%FIXME: make better references to the tables
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with no epenthetic vowel}\label{exrnnovowel}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			%FIXME: should the glosses here and below be morpheme-by-morpheme for consistency, or is a more colloquial gloss okay?
			/мурн-(S)I/ & [мурду] & \eng{his/her/its nose} \\
			/мурн-Iм-DAн/ & [мурдуман] & \eng{from my nose} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{0.5em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн-(S)I/ & [орду] & \eng{his/her/its place} \\
			/орн-Iм-DAн/ & [ордуман] & \eng{from my place} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}

>>>>

Kyrgyz has a number of nouns that underlyingly end with a consonant cluster consisting of two liquids, but which are split by an epenthetic vowel syllable-finally, as shown in table~\ref{exrnvowel}.
\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with epenthetic vowel}\label{exrnvowel}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/мурн/ & [мурун] & \eng{nose} \\
			/мурн-LAр/ & [мурундар] & \eng{nose--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{2em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн/ & [орун] & \eng{place} \\
			/орн-LAр/ & [орундар] & \eng{places--\gmk{PL}} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}
However, when a vowel-initial suffix follows the noun, as in table~\ref{exrnnovowel}, the epenthetic vowel is absent, and rules of consonant unfaithfulness involving consonant clusters (e.g., desonorisation) become relevant as well.

\begin{table}[htbp]
	\caption{Examples of /рн/-final nouns surfacing with no epenthetic vowel}\label{exrnnovowel}
\begin{center}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			%FIXME: should the glosses here and below be morpheme-by-morpheme for consistency, or is a more colloquial gloss okay?
			/мурн-(S)I/ & [мурду] & \eng{his/her/its nose} \\
			/мурн-Iм-DAн/ & [мурдуман] & \eng{from my nose} \\
			\bottomrule
		\end{tabular}
	}
	\hspace{0.5em}
	\subtable{
		\begin{tabular}{lll}
			\toprule
			underlying & surface & gloss \\
			\midrule
			/орн-(S)I/ & [орду] & \eng{his/her/its place} \\
			/орн-Iм-DAн/ & [ордуман] & \eng{from my place} \\
			\bottomrule
		\end{tabular}
	}
\end{center}
\end{table}
